---
title: "Hello? Anyone There?"
subtitle: "A Call Centre Simulation"
author: "Sean"
date: '2019-12-17'
image: ''
bigimg:
- desc: Photo by Moose Photos from Pexels
  src: /img/angry_phone.jpg
slug: call-center-simulation
tags:
- order statistics
- simulation
- queueing theory
categories: R
type: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

day_rate <-  1/3
night_rate <-  1/7
num_day <- 3
num_night <- 6
call_cost_day <- 31.85
call_cost_night <-  27.94

```

#TODO: add in single observational call to budget simulation! current results reflext your call being answered not your problem being resolved

This post will deal with how to approach a problem using simulation. Furthermore, some theory on the distribution of order statistics shall be touched on as well as some very basic Queueing Theory concepts.

# Context

Say you are trying to reach an international number. It can cost a fortune to reach an operator on the other side of the world. You have limited funds and you wish to simulate a call center system to estimate how likely you are to reach an operator before you run out of money.

# Assumptions

It is critical to be extremely clear about the assumptions you are making when simulating a system. It should be clear in the analyst's mind as to what the limitations are of the simulation. A simulation needn't be overly complex, but it should be sophisticated enough to capture the core dynamics that are being investigated.

* Presume that it is a 24/7 line , but the number of call operators differs each shift. 

* Assume that the call operators' call resolution times are *homogenous* within a given shift but each shift has a different mean waiting time. For example, on average you may need to wait longer when calling at night.

* Assume that people who are also waiting for an operator do not leave the queue. They wait until their calls are answered. This refers to the concept of *reneging*. More on this can be found by searching for 

* We will also assume a 'memoryless' property to waiting in line. What this means is that if you have already been waiting $T$ minutes (say 10 minutes in line), then this does **not** affect your chances of your call being picked up within the next *t* minutes. Simply put, the amount of time you have already been waiting does not influence your chances at all of your call being picked up.

put memoryless property ehre?

<!-- Question: What strategy should you employ to *maximise the probability of reaching a call operator* given you don't use up all of your airtime? -->


# Distributional Assumptions

A reasonable distribution to model the time a *SINGLE call operator* spends on an individual phone call is the Exponential distribution.

* Memoryless process: the Exponential distribution has this property discussed earlier.

* Independence between call operators.

* Exponential distribution is continuous and defined for positive variables.

Let there be two shifts per day: Day (8 am to 8 pm) and Night (8pm to 8am). I shall let the Day shift operators be more effective and thus they will have a lower mean call resolution time.

$\lambda_{Operator}$ is the rate parameter which characterises the Exponential distribution. The average time a call operator will spend on a call in *minutes* is $\frac{1}{\lambda_{Operator}}$.

# Simulation Hyperparameters

Taking into account the previous two sections, we can set hyperparamers which will characterise the nature of the simulation

* Number of Day operators: 3
* Number of Night operators: 6
* Each call resolution time for a single Day shift call $t_{Day} \sim E(\lambda_{day} = 1/3)$
* Each call resolution time for a single Night shift call $t_{night} \sim E(\lambda_{night} = 1/7)$
* The cost per minute for calling an international number during the day is R31.85 and at night it is R27.94.

The respective distributions are plotted below:

```{r call_operators}


#Simulate resolution times for a single operator
day_simulations <- rexp(n = 10000, rate = day_rate)
night_simulations <- rexp(n = 10000, rate = night_rate)
#Day and Night plots
      density(day_simulations) %>% 
      plot(xlab = "Call Resolution Time in minutes", main = "Comparison of Call Resolution Times for a single operator",xlim = c(0,50),ylim = c(0,0.4), col = "blue")
lines(density(night_simulations), col = "red")
legend("topright", c("Day Shift","Night Shift") ,col = c("blue","red"), lty = 1)
      
    

```

Variance in call resolution times at night is much higher than during the day. Since on average, the day time staff resolve calls more quickly, fewer staff are kept during this shift. Less experienced staff are kept on during the night shift.

**Should you call at night when the international call fees are lower with more staff on duty or should you call during the day when the calls are more expensive but the fewer staff on duty are more efficient?**

<!-- Since the daytime staff are more efficient, fewer staff are on duty. Let this value be `r paste0(num_day)`. There are more nighttime staff at `r paste0(num_night)`. -->

# Simulating Waiting Times

Using the Exponential information for each call operator, a distribution for the amount of time that *you* would need to wait can be derived through simulation.

A function needs to be written to take into account:

* The number of call operators
* The call resolution rate of these operators

This function will should return the *minimum* call resolution times for the N number of call operators. This represents the time you would have to wait if you were to stay on the line until someone was free to take your call. This shall be referred to as the "First Available" times.

```{r}

call_pickup <- function(num_operators, pickup_rate){
  
  pickup_time <- rexp(n = num_operators,
                      rate = pickup_rate) %>% 
                      min()
  return(pickup_time)
}


```

Let's simulate the First Available waiting times for 10000 calls at each shift. What would be the distribution of waiting times?

```{r}

par(mfrow = c(1,2))

#Simulate the waiting times for each shift
daytime_first_sim <-
  replicate(n = 10000,
            expr = call_pickup(num_operators = num_day, pickup_rate = day_rate))

night_first_sim <-
  replicate(n = 10000,
            expr = call_pickup(num_operators = num_night, pickup_rate = night_rate))

par(mfrow = c(2,1))
density(daytime_first_sim) %>% 
  plot(xlab = "Time Spent Waiting for First Available Operator", main = 'First Available Waiting Times',xlim = c(0,50),ylim = c(0,0.85), col = "blue")
lines(density(night_first_sim),col = "red")


density(day_simulations) %>% 
      plot(xlab = "Call Resolution Time in minutes", main = "Single Operator Resolution Times",xlim = c(0,50),ylim = c(0,0.4), col = "blue")
lines(density(night_simulations), col = "red")
legend("topright", c("Day Shift","Night Shift") ,col = c("blue","red"), lty = 1)
      
    
```

Not surprising in that the minimum of Exponential variables is indeed an exponential distribution. We can see that the First Available waiting times are MUCH less variable now because you are waiting for the first available operator out of many. The fact that there are 3 fewer operators during the day shift compared to the night is more of a factor now. 

# Analytically Deriving Waiting Times

Not surprising in that is indeed an exponential distribution. We can see that the waiting times are MUCH less variable now because you are waiting for the first available operator out of many. The median waiting times are also lower. This is the same as simply calculating the distribution of the *Minimum* or rank 1 statistic for a vector of random variables.

In fact, this could be solved analytically using some basic probability theory surrounding cumulative density functions (CDFs). The thought process is to derive a distributional expression of the probability of observing a specific $x^*$ as being the minimum value out of a sample of N data points.

$\text{Cumulative Density Function(CDF)} = F(x^*) = Pr(X\leq x^*)$

For it to be the minimum, we consider probability of a value being greater than $x^*$

$Pr(X\geq x^*) = 1 - F(x^*)$

The above expression is the probability in which a random variable X has the same value or greater than $x^*$. For $x^*$ to be the minimum value from a sample draw of N random variables, each and every value needs to be greater than or equal to $x^*$.

Probabilisticly, assuming independence:

$Pr(X_1, X_2 ... X_N \geq x^*)  = (1 - F(x^*))^N$

$\text{In order to make the above expression into a CDF for the minimum:}$

$F(x_{min}) = Pr(X\leq x^*)  = 1 - (1 - F(x^*))^N$

We are nearly there. The last equation above shows the **general expression** for the CDF of the minimum order statistic out of a sample of $N$ data points. Notice how no distributional assumptions were made in deriving this. In our case, the CDF for the Exponential distribution is substituted where $F(x^*)$ stands. If we were dealing with Poisson variables for example, we would just need to substitute $F(x^*)$ with the CDF of the Poisson distribution.

In order to achieve the density function, we would simply differentiate the expression with respect to the random variable $x$.

The density distribution for the Available Waiting times is thus:

$t^{(1)}_{day} \sim \lambda_{day}^{2} N\left(-e^{\lambda_{day}(-t^{(1)}_{day})}\right)\left(1-\lambda_{day} e^{\lambda_{day}(-t^{(1)}_{day})}\right)^{N-1}$

We won't need to use the density distribution above as we can simulate variables from the CDF using the Probabibility Inverse Transform method (PIT) method.

I shall skip over the explanation, but essentially variables from a desired distribution can be simulated directly through the use of uniform random variables

As a sanity check, the analytical and simulated First Available waiting times are compared to show that we have indeed solved the problem analytically.

```{r first_order}

first_order_sim <- function(n,rate){
  u <- runif(n =1)
  #Probability Inverse Transform of the CDF of the Min 
  -log(1-u)/(n*rate)
}

analytical_first_order <- replicate(n = 10000,expr =  first_order_sim(n = num_day,rate = day_rate)) 

par(mfrow = c(1,1))
density(daytime_first_sim) %>% 
  plot(xlab = "Time Spent Waiting for First Available Operator", main = 'First Available Waiting Times: Day',xlim = c(0,10),ylim = c(0,0.85), col = "blue")
lines(density(analytical_first_order),lty = 4, col = "black")
legend("topright", c("Simulated","Analytical"), col =c("blue","black"), lty = c(1,4))
```

# Will Anyone Answer Me?

We have an explicit analytical set of solutions for the First Available times. As a remidner, our central question is to answer: 

**Should you call at night when the international call fees are lower with more staff on duty or should you call during the day when the calls are more expensive but the fewer staff on duty are more efficient?**

Under a fixed budget, let's say R50, should we phone at night or during the day? I shall simulate the scenario 10000 times and caluclate the success rates. To simplify matters, only two strategies are conisdered: phoning during the day or phoning at night.

```{r call_scenario, echo = TRUE}

call_scenario <-
  function(num, #number of  operators
           rate, #lambda characterising their performance
           budget = 50, #How much airtime you have to spend
           call_cost = 31.85, #cost per MINUTE
           period = 0 #0 is day and 1 is night
           ){

        #Simulate When An Operator Will Be Available
        waiting_time <- first_order_sim(n = num,rate = rate)
        #Calculate Money Spent Waiting for Call to be answered
        money_spent <- waiting_time * call_cost
        budget <- max(0, budget - money_spent)
        
        if(budget > 0){
          #IF you have money leftover and You didn't hang up, Call was Answered
          return(c(result = TRUE,period = period ))
        }else{
          return(c(result = FALSE, period  = period))
        }
        
}



```



```{r illustrative_sim, eval = F}

 
simulation_results_day <- replicate(n = 10000,
         expr = call_scenario(num = num_day,
                              rate = day_rate,
                              budget = 50,
                              call_cost = call_cost_day
                         )) %>% t()

simulation_results_night <- replicate(n = 10000,
         expr = call_scenario(num = num_night,
                              rate = night_rate,
                              budget = 50,
                              period = 1,
                              call_cost = call_cost_night
                         )) %>% t()


simulation_results_day[,1] %>% mean()
simulation_results_night[,1] %>% mean()


```







<!-- This isn't that interesting. What we want is to develop the **optimal** calling strategy here. An added complication is that the proportion of successful calling scenarios is a stochastic output. The same strategy will output different success rates. -->

<!-- To illustrate this stochastic nature, the usual optim() function shall be used to find the optimal calling proportion strategy for the scenario above. This solution shall be solved 1000 times. -->



<!-- ```{r loss_function, eval = FALSE} -->
<!-- #Define a loss function to output success rate  -->


<!-- loss_fn <- function(prop_day_budget, -->
<!--                     sim_iterations = 1000, -->
<!--                     num_day, -->
<!--                     day_rate, -->
<!--                     num_night, -->
<!--                     night_rate, -->
<!--                     total_budget -->
<!--                                 ){ -->



<!--   simulation_results <-     replicate(n = sim_iterations, -->
<!--                                       expr = call_scenario( -->
<!--                                                       num_day, -->
<!--                                                       day_rate, -->
<!--                                                       num_night, -->
<!--                                                       night_rate, -->
<!--                                                       total_budget -->
<!--                                                       ) -->
<!--                                       ) %>% t() -->

<!--   return(mean(simulation_results[,1])) -->

<!-- } -->


<!-- # loss_fn(prop_day_budget = 0.25, -->
<!-- #                     num_day = 6, -->
<!-- #                     day_rate = 1/10, -->
<!-- #                     num_night = 1/15, -->
<!-- #                     night_rate = 3, -->
<!-- #                     total_budget = 50) -->


<!-- solution <- c() -->
<!-- solution_success <- c() -->

<!-- for(i in 1:500){ -->

<!--   non_stochastic_optim <- optim(par = 0.5, -->
<!--                               lower = 0, -->
<!--                               upper = 1, -->
<!--                                 fn = loss_fn, -->
<!--                                 num_day = 6, -->
<!--                                 day_rate = 1/10, -->
<!--                                 num_night = 15, -->
<!--                                 night_rate = 1/30, -->
<!--                                 total_budget = 50, -->
<!--                               method = "Brent") -->

<!--   solution[i] <- non_stochastic_optim$par -->
<!--   solution_success[i] <- non_stochastic_optim$value -->

<!-- } -->

<!-- plot(density(solution),xlab ="Optimal Prop Day Time Calling Solution", main ="Solving the problem with optim 100 Times") -->

<!-- plot(density(solution_success), main = "Distribution of Strategy Success Rates")         -->


<!-- plot(x = solution , y = solution_success, main = "Optimal Prop vs Solution Success Achieved") -->

<!-- ``` -->

<!-- It can be seen that the loss function output is a Gaussian random variable. However, the hyperparameter, the proportion of the budget to spend during the day, appears to be a Gaussian mixture distribution -->
